---
title: "CodeBook"
output: html_notebook
---

Initially, the feature labels are stored inside a variable called features:

```{r}
features <- read.table("./UCI HAR Dataset/features.txt")[,2]
```

Both test and train sets are then stored inside two variables named test_df1 and train_df1. The labels are stored in test_df2 and train_df2. The subjects are stored in the test_sub and train_sub variables:

```{r}
train_df1 <- read.table("./UCI HAR Dataset/train/X_train.txt",col.names = features)
train_df2 <- read.table("./UCI HAR Dataset/train/y_train.txt",col.names = c("Activity"))
train_sub <- read.table("./UCI HAR Dataset/train/subject_train.txt",col.names = c("Subject"))

test_df1 <- read.table("./UCI HAR Dataset/test/X_test.txt",col.names = features)
test_df2 <- read.table("./UCI HAR Dataset/test/y_test.txt",col.names = c("Activity"))
test_sub <- read.table("./UCI HAR Dataset/test/subject_test.txt",col.names = c("Subject"));
```

Then the labels, sets and subjects are merged and stored in variables called test_df3 and train_df3. After merging, only the variables containing the words "mean" and "std" are selected:

```{r}
train_df3 <- cbind(train_sub,train_df2,train_df1)
test_df3 <- cbind(test_sub,test_df2,test_df1)

train_df3 <- train_df3 %>% select(Subject,Activity,contains("std"), contains("mean"))
test_df3 <- test_df3 %>% select(Subject,Activity,contains("mean"), contains("std"))
```

Finally both data sets are merged and stored in a variable called final_df. We arrange it by Subject and Activity and set both these variables to factors:

```{r}
final_df <- rbind(train_df3,test_df3)
final_df <- arrange(final_df,Subject, Activity)
final_df$Subject <- as.factor(final_df$Subject)
final_df$Activity <- as.factor(final_df$Activity)
```

The levels of the Activity variable are assigned descriptive values, other variable names are assigned descriptive names:

```{r}
levels(final_df$Activity) <- c("WALKING","WALKING UPSTAIRS","WALKING DOWNSTAIRS","SITTING","STANDING","LAYING")
names(final_df) <- gsub("^t","Time",names(final_df))
names(final_df) <- gsub("^f","Frequency",names(final_df))
names(final_df) <- gsub("Acc","Accelerometer",names(final_df))
names(final_df) <- gsub("Gyro","Gyroscope",names(final_df))
names(final_df) <- gsub("Mag","Magnitude",names(final_df))
names(final_df) <- gsub("BodyBody","Body",names(final_df))
names(final_df) <- gsub("tBody","TimeBody",names(final_df))
names(final_df) <- gsub("angle","Angle",names(final_df))
names(final_df) <- gsub("gravity","Gravity",names(final_df))
names(final_df) <- gsub("-mean()",".Mean",names(final_df))
names(final_df) <- gsub("-std()",".STD",names(final_df))
names(final_df) <- gsub("-freq()",".Frequency",names(final_df))
```

In the end, we summarize the data to obtain the mean of all variables grouped by Subject and Activity:

```{r}
summarized_df <- final_df %>% group_by(Subject,Activity) %>% summarise(across(everything(), mean))
```